fuzzy-matching.Rmd

### 2019-02-19 notebook/algae.Rmd
improving clean names utility
update example, docs

```{r}
taxadb::td_create() # only needed once.
algae <- read_csv("~/projects/algae-names/algae_uncleanednames_NAdropped.csv")
```

@@

```{r}
length(ott$input)
length(unique(ott$input))
```

@@

Wow, so some of our entries have a whole lot more words than a species name: the top 6 entries have dozens of words:

```{r}
head(wordcount$input)
```

many more (1,551) have 4 words (three spaces), indicating subspecies and varities:

@@

```{r}
clean_unmatched <- unmatched %>% clean_names() %>% unique()
length(clean_unmatched)
```
This gives us 9303 unmatched names to resolve.

@@

With these clean names, another 1,927 can be matched to OTT ids, leaving 7,526 unmatched:

```{r}
length(unmatched2)
head(unmatched2, 10)
```

@@

```{r}
```

>>>>

```{r}
taxadb::td_create("all") # only needed once.
algae <- read_csv("~/projects/algae-names/algae_uncleanednames_NAdropped.csv")
```

@@

```{r}
length(ott_ids$input)
length(unique(ott_ids$input))
```

@@

Wow, so some of our entries have a whole lot more words than a species name: the top 6 entries have dozens of words.  Many more (1,551) have 4 words (three spaces), indicating subspecies and varities:

@@

```{r}
clean_unmatched <- unmatched %>% clean_names() %>% unique()
length(clean_unmatched)
```
This gives us `r length(clean_unmatched)` unmatched names to resolve.

@@

With these clean names, another `r length(clean_unmatched) - length(umatched2)` can be matched to OTT ids, leaving `r length(umatched2)` unmatched:

```{r}
length(unmatched2)
head(unmatched2, 10)

ott <- taxa_tbl("ott") %>% collect()
ott_mash <- ott %>% mutate(input = clean_names(scientificName))
```

@@

```{r}
gbif_ids <- unmatched2 %>% ids("gbif")
unmatched3 <- gbif_ids %>% filter(is.na(acceptedNameUsageID)) %>% pull(input)
length(names(umatched3))
```

FIXME also try matching against `clean_names` versions of scientificName column on the database side; in particular, on the synonyms.
